---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**
- 40% more than the mean because there are far more color options than just red or black
- The M&M's are randomly supplied which means that different groups are going to get different amounts and varying amount of each color.
- The extremes would balance out and form a more normal looking distrubution.
- Every color of M&M represents a different answer choice on the survey. At the end there will be a random distribution of survey responses.

------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**
The sample uses 23% of the total UMD population. The sample is made up of only 38% men while the campus population is about 50%. 55% of the sample was women compared to 50% of the campus population being women. One thing that rasies questions was the fact that international students were used for the graduate student sample. The sample is not distributed in a way that is representative of the entire population. 
------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
rows_data <- tibble(age = sample(18:80, 1000, replace = TRUE))
```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_100 <- rows_data |> sample_n(100)
```

### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_100 |> summarize(mean = mean(age), sd = sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
rows_data |> summarize(mean = mean(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics
**A** 

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

**A**: The mean of the sample is lower than the mean of the whole population. The mean is 1.1 higher in the total population statistics. The standard deviations are close however the total population has a lower standard deviation than the sample. For smaller sample sizes the mean and standard deviation stay relatively the same.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
