---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

ANSWER HERE: The main thing that stands out to me from a reporting standpoint is the gap between the highest paid person and the second highest paid person. Michelle Barrans is making significantly more than the next highest paid person, Farah Ahmad. Why is Michelle making so much compared to everyone else? This value will certainly cause problems for analysis becasue it is an outlier. Also, the 11 people that do not have a salary will cause problems when analyzing the data.  

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |>
  summarize(mean_salaries=mean(salary),
            median_salaries=median(salary))

```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  summarize(mean_salaries=mean(salary),
            median_salaries=median(salary),
            min_salaries=min(salary),
            max_salaries=max(salary))

```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

ANSWER HERE: The minimun and maximum values tell us that there are absolutely outliers in the data. Both 0 and 251,258 are much lower and higher than their closest salary comparison. While the mean and median aren't too far off, there it still a distinct difference. The outliers will skew the mean a bit, while the median is probably more accurate.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

ANSWER HERE: I think the story here is why is Michelle making so much more money than anyone one else? What does she do that is so much more important that the other employees. The next steps would be to figure out what she does. That could mean reaching out to her or her department to get a better understanding of her day-to-day tasks.
