---
title: "HW7_MandMs"
date: 10/1/2025
---


In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load an M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
candy <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
set.seed(1234) 
sample_5 <- candy |> sample_n(5)
```

```{r}
sample_10 <- candy |> sample_n(10)
```

### Task 4: Calculate the mean, median and standard deviation of red M&Ms from each of the samples.
```{r}
candy_sample <- sample_5 |>
  summarize(mean = mean(red), median = median(red), sd = sd(red))
```

```{r}
candy_sample2 <- sample_10 |>
  summarize(mean = mean(red), median = median(red), sd = sd(red))
```
### Task 5: Calculate the mean, median and standard deviation of red M&Ms from the population

```{r}
candy_population <- candy |>
  summarize(mean = mean(red), median = median(red), sd = sd(red))
```

### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Include a headline. Use a consistent binwidth.

```{r}
sample_5 |>
 ggplot() + 
  geom_histogram(aes(x = red), binwidth = 2) 
```
```{r}
sample_10 |>
 ggplot() + 
  geom_histogram(aes(x = red), binwidth = 2) 
```
```{r}
candy |>
 ggplot() + 
  geom_histogram(aes(x = red), binwidth = 2) 
```
### Task 7: Answer questions


Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples? : Each have similar means and medians. I would use mean and standard deviation for understanding the accuracy.

What does this exercise tell you about the relationship between sample size and estimate accuracy? : For the most part, the smaller the sample size, the less accurate the estimates will be. However, in this sample the estimates are quite close.

How does random sampling introduce variation into statistical estimates? : With random sampling, it is possible that pieces of data from the same part of the population can end up in the same sample, giving statistical variation in the estimates.

If you repeated this process, would the samples always be identical? Why or why not? : The samples would not always be identical. The data set is large enough where getting identical samples would be extremely unlikely.


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
